Urbanization and Public Transportation of Los Angeles

### 2023-11-09 Paper.Rmd
revert julian test
Revert "Initial population data"
This reverts commit 6d8f0782ef3a20d55b434ed1a540f95499aec9f5.

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(ggplot2)
library(latex2exp)
library(dplyr)
library(tidyr)
library(scales)
library(stargazer)
ridership_data <- read.csv("data/April_Transit_Ridership.csv")
population_data <- read.csv("data/Los_Angeles_Population.csv")
budget_data <- read.csv("data/City_Budget_Expenditures_on_Transportation_20231106.csv")
```

@@

Testhfgg

@@

# Predict the rate of change for the fitted polynomial models
ridership_data$PredictedBusROC_poly <- predict(bus_model_poly, newdata=ridership_data)
ridership_data$PredictedRailROC_poly <- predict(rail_model_poly, newdata=ridership_data)

# Normalize the ridership data with the population data to get ridership per capita
ridership_data <- ridership_data %>%
  left_join(population_data, by = "Year") %>%
  mutate(
    RidershipPerCapitaBus = Total.Monthly.Bus / population,
    RidershipPerCapitaRail = Total.Monthly.Rail / population
  )

# Calculate the rate of change for Ridership per Capita
ridership_data$BusRateOfChangePerCapita <- c(NA, diff(ridership_data$RidershipPerCapitaBus))
ridership_data$RailRateOfChangePerCapita <- c(NA, diff(ridership_data$RidershipPerCapitaRail))

# Linear and polynomial regression for ridership per capita
bus_model_per_capita <- lm(BusRateOfChangePerCapita ~ Year, data = ridership_data)
rail_model_per_capita <- lm(RailRateOfChangePerCapita ~ Year, data = ridership_data)
```

@@

```{r predict-trend-usage}
# Predict the rate of change per capita for all years using the linear and polynomial models
# (This assumes you have a model fitted to per capita data)
all_years <- data.frame(Year = seq(min(ridership_data$Year), 2027))
all_years <- left_join(all_years, population_data, by = "Year")

all_years$PredictedBusROCPerCapita <- predict(bus_model_per_capita, newdata=all_years)
all_years$PredictedRailROCPerCapita <- predict(rail_model_per_capita, newdata=all_years)

# Assuming the rate of change per capita is per year, integrate the rate of change to get the total ridership per capita
first_known_bus_ridership_per_capita <- head(ridership_data$RidershipPerCapitaBus, 1)
first_known_rail_ridership_per_capita <- head(ridership_data$RidershipPerCapitaRail, 1)

all_years$RidershipPerCapitaBus <- first_known_bus_ridership_per_capita + cumsum(c(0, all_years$PredictedBusROCPerCapita[-1]))
all_years$RidershipPerCapitaRail <- first_known_rail_ridership_per_capita + cumsum(c(0, all_years$PredictedRailROCPerCapita[-1]))
```

```{r format-and-combine-data}
# Format and Combine Data
# Convert the predicted data to long format
predicted_data_long <- all_years %>%
  select(Year, RidershipPerCapitaBus, RidershipPerCapitaRail) %>%
  pivot_longer(cols = c("RidershipPerCapitaBus", "RidershipPerCapitaRail"),
               names_to = "Type",
               values_to = "TotalRidership") %>%
  mutate(Prediction = "Predicted")

# Now, make sure the actual data is in the same long format with the same column names
actual_data_long <- ridership_data %>%
  select(Year, RidershipPerCapitaBus, RidershipPerCapitaRail) %>%
  pivot_longer(cols = c("RidershipPerCapitaBus", "RidershipPerCapitaRail"),
               names_to = "Type",
               values_to = "TotalRidership") %>%
  mutate(Prediction = "Actual")

@@

# Filter out the extreme predicted values for the polynomial bus data
combined_data_long <- combined_data_long %>%
  filter(!(Type == "RidershipPerCapitaBus" & TotalRidership < -10000000 & Prediction == "Predicted"))

@@

```{r plot-trend-usage, fig.cap="Line plot of public transportation ridership over time with regression lines"}
# Create the line plot with observed and predicted per capita data
p <- ggplot(combined_data_long, aes(x = Year, y = RidershipPerCapita, color = Type)) +
  geom_line(aes(linetype = Prediction)) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(title = "Trend in Public Transportation Usage Per Capita",
       x = "April of Each Year",
       y = "Ridership Per Capita",
       color = "Transportation Type",
       linetype = "Data Type") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10), limits = c(min(ridership_data$Year), max(all_years$Year)))

# Print the plot
print(p)

# Sources:

- Transit ridership data: Los Angeles Metro "L.A. METRO TRANSIT RIDERSHIP UP 10 PERCENT, SETS POST-PANDEMIC RECORD", Patrick Chandler

- Los Angeles Population Data (real and predicted): WorldPopulationReview "Los Angeles, California Population"

>>>>

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(ggplot2)
library(latex2exp)
library(dplyr)
library(tidyr)
library(scales)
library(stargazer)
ridership_data <- read.csv("data/April_Transit_Ridership.csv")
budget_data <- read.csv("data/City_Budget_Expenditures_on_Transportation_20231106.csv")
```

@@

@@

# Predict the rate of change for the fitted polynomial models
ridership_data$PredictedBusROC_poly <- predict(bus_model_poly, newdata=ridership_data)
ridership_data$PredictedRailROC_poly <- predict(rail_model_poly, newdata=ridership_data)
```

@@

```{r predict-trend-usage}
# Predict Trend Usage
# Extend the data frame to include all years for prediction
all_years <- data.frame(Year = seq(min(ridership_data$Year), 2027))

# Predict the rate of change for all years using the linear and polynomial models
all_years$PredictedBusROC <- predict(bus_model, newdata=all_years)
all_years$PredictedRailROC <- predict(rail_model, newdata=all_years)
all_years$PredictedBusROC_poly <- predict(bus_model_poly, newdata=all_years)
all_years$PredictedRailROC_poly <- predict(rail_model_poly, newdata=all_years)

# Initialize the first known ridership value
first_known_bus_ridership <- head(ridership_data$Total.Monthly.Bus, 1)
first_known_rail_ridership <- head(ridership_data$Total.Monthly.Rail, 1)

# Assuming the rate of change is per year, integrate the rate of change to get the total ridership
# We start from the first known ridership value and accumulate the predicted rate of change
all_years$Total.Monthly.Bus <- first_known_bus_ridership + cumsum(c(0, all_years$PredictedBusROC[-1]))
all_years$Total.Monthly.Rail <- first_known_rail_ridership + cumsum(c(0, all_years$PredictedRailROC[-1]))
all_years$Total.Monthly.Bus_poly <- first_known_bus_ridership + cumsum(c(0, all_years$PredictedBusROC_poly[-1]))
all_years$Total.Monthly.Rail_poly <- first_known_rail_ridership + cumsum(c(0, all_years$PredictedRailROC_poly[-1]))
```

```{r format-and-combine-data}
# Format and Combine Data
# Convert the predicted data to long format
predicted_data_long <- all_years %>%
  select(Year, Total.Monthly.Bus, Total.Monthly.Rail, Total.Monthly.Bus_poly, Total.Monthly.Rail_poly) %>%
  pivot_longer(cols = c("Total.Monthly.Bus", "Total.Monthly.Rail", "Total.Monthly.Bus_poly", "Total.Monthly.Rail_poly"),
               names_to = "Type",
               values_to = "TotalRidership") %>%
  mutate(Prediction = "Predicted")

# Now, make sure the actual data is in the same long format with the same column names
actual_data_long <- ridership_data %>%
  select(Year, Total.Monthly.Bus, Total.Monthly.Rail) %>%
  pivot_longer(cols = c("Total.Monthly.Bus", "Total.Monthly.Rail"),
               names_to = "Type",
               values_to = "TotalRidership") %>%
  mutate(Prediction = "Actual")

@@

# Filter out the extreme predicted values for the polynomial bus data
combined_data_long <- combined_data_long %>%
  filter(!(Type == "Total.Monthly.Bus_poly" & TotalRidership < -10000000 & Prediction == "Predicted"))

@@

```{r plot-trend-usage, fig.cap="Line plot of public transportation ridership over time with regression lines"}
# Plot Trend Usage
# Create the line plot with observed data
p <- ggplot(combined_data_long, aes(x = Year, y = TotalRidership, color = Type)) +
  geom_line(aes(linetype = Prediction)) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(title = "Trend in Public Transportation Usage",
       x = "April of Each Year",
       y = "Total Monthly Ridership",
       color = "Transportation Type",
       linetype = "Data Type") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10), limits = c(min(ridership_data$Year), max(all_years$Year))) +
  scale_y_continuous(labels = scales::label_comma(), breaks = scales::pretty_breaks(n = 14))

# Add predicted lines for past years (where we have actual data)
predicted_data_long <- all_years %>%
  select(Year, Total.Monthly.Bus, Total.Monthly.Rail, Total.Monthly.Bus_poly, Total.Monthly.Rail_poly) %>%
  pivot_longer(cols = c("Total.Monthly.Bus", "Total.Monthly.Rail", "Total.Monthly.Bus_poly", "Total.Monthly.Rail_poly"),
               names_to = "Type",
               values_to = "TotalRidership") %>%
  mutate(Prediction = "Predicted") %>%
  filter(Year <= max(ridership_data$Year))

# Overlay this on the plot with a different line type
p <- p +
  geom_line(data = predicted_data_long, aes(x = Year, y = TotalRidership, color = Type, linetype = Prediction))

print(p)
```
